---
title: 'Reproducible Research: Peer Assessment 1'
output: 
        html_document:
                keep_md: true
---

## Loading and preprocessing the data
1. As the data is in our working directory we will read it and preprocess it.  
```{r preprocess}
unzip("activity.zip")
data <- read.csv("activity.csv")
```  

## What is mean total number of steps taken per day?
Calculate the total number of steps per day  
```{r daymean step}  
library(dplyr)
dataDate <- group_by(data, date)
sum <- summarize(dataDate, steps=sum(steps, na.rm=TRUE))
``` 

2. Histogram of the total number of steps per day
```{r histogram steps}  
library(ggplot2)  
with(sum, qplot(steps, binwidth=2500, main="Total number of steps per day"))  
```

3. Mean and median of total number of steps per day  
```{r steps}
summary(sum$steps)
```

## What is the average daily activity pattern?
4. Time series plot of the average steps for 5 min interval for each day. 
```{r daily activity}
mean <- summarize(dataDate, steps=mean(steps, na.rm=TRUE))
mean$date <- as.Date(as.character(mean$date))
with(mean, plot(date, steps, main="average daily 5 min interval steps over time"))
with(mean, lines(date, steps))
```  

5. Maximum number of steps, by daily average :   
```{r steps max}  
filter(mean, steps==max(steps, na.rm=TRUE))  
```  

## Imputing missing values
6. Number of missing values in the data, will appear in the TRUE column  
```{r NAs}
table(!complete.cases(data))
```  

Function to replace the NA with the mean of steps per day, and create a new dataset : cleanData with no NAs.  
```{r functionNA}  
cleanData <- dataDate
temp <- mean(mean$steps, na.rm=T)
cleanData[is.na(cleanData$steps),1] <- temp
```  

7. Histogram of the total number of steps taken each day  
```{r total steps}
cleanSum <- summarize(cleanData, steps=sum(steps, na.rm=TRUE))
with(cleanSum, qplot(steps, binwidth=2500, main="Total number of steps per day"))
```   
  
The plot differs, we don't have the peak on the 0.  

Mean and median of the total steps.  
```{r steps summary}
summary(cleanSum$steps)
```  
The median and the mean differs slightly.  

## Are there differences in activity patterns between weekdays and weekends?
8. Create the factor variable wDays and plot the mean steps per interval for each day in a two panel weekday vs weekend plot.  
```{r}
cleanData$day <- weekdays(as.Date(as.character(cleanData$date)))
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
cleanData$wDay <- factor(cleanData$day %in% weekdays, levels=c(TRUE, FALSE), labels=c("weekday", "weekend"))
dData <- group_by(cleanData, wDay, date)
mean2 <- summarize(dData, steps=mean(steps))
mean2$date <- as.Date(as.character(mean2$date))
library(lattice)
xyplot(steps~date|wDay, mean2, layout=c(1,2), type="l")
```